New lower bounds for Schur and weak Schur numbers

### 2021-05-02 article/schur.tex
modfis demos

\begin{proof}[\textsc{Proof.}]
We denote by \(f\) the colouring associated to the \(SF\)-template of lenght \(p\)
and \(g\) the one associated to the sum-free partition of \([\![1,q]\!]\).

\[ f : [\![1,p]\!] \longrightarrow [\![1,k+1]\!] \text{ and } \forall (x,y) \in [\![1,p]\!]^2, \left\{
\begin{array}{ll}
	f(x) = f(y) \leqslant k \\
	x + y > p
\end{array}
\right.
\Longrightarrow f(x+y-p) \neq f(x)
\]
with also the sum-free condition : \[ \forall (x,y) \in [\![1,p]\!]^2 \text{, } f(x) = f(y)
\Longrightarrow f(x+y) \neq f(x) \]
Moreover, by definition of \(m_{k+1}\), if \(x < m_{k+1}\), \(f(x) \leqslant k\).

\[g : [\![1,q]\!] \longrightarrow [\![1,n]\!] \text{ and } \forall (x,y) \in [\![1,q]\!]^2 \text{, } g(x) = g(y)
\Longrightarrow g(x+y) \neq g(x)
\]

We now define \(h : [\![1,pq+m_{k+1}-1]\!] \longrightarrow [\![1,n+k]\!] \) as follows :
\( \forall x \in [\![1,pq + m_{k+1}-1]\!] \), we write \(x = \alpha p - u\) where \(\alpha \in [\![1,q+1]\!] \) and \( u \in
[\![0,p-1]\!] \).
This decomposition if of course unique. \\

If \(f(p-u) \leqslant k\), we set \(h(x) = f(p-u)\), else (i.e \(f(p-u) = k + 1)\) \(h(x) = k + g(\alpha) \). \\
If \(\alpha = q+1\), we have a problem plugging it into \(g\). Hopefully, this issue never arises : if 
\(\alpha = q+1\), we have \(u > p - m_{k+1}\) since \(x \in [\![1,pq + m_{k+1}-1]\!]\), hence \(p-u < m_{k+1}\) and \(f(p-u) \leqslant k\). \\

Prooving that the partition of \([\![1,pq]\!]\) induced by \(h\) is sum-free will complete the proof.\\

Let \(x,y \in [\![1,pq + m_{k+1}-1]\!]\), with \(h(x) = h(y)\) and \(x+y \leqslant pq+m_{k+1}-1\). We write \(x = \alpha p - u\) and \(y =
\beta p - v\). We simply need \(h(x+y) \neq h(x)\). \\

We first examine the case where \(h(x) \leqslant k\). We assume \(h(x+y) \leqslant k\), otherwise we are fine. 
Thus by definition of \(h\), we have \(h(x) = f(p-u) = f(p-v) = h(y)\).
\begin{itemize}
\item If \(u+v < p\), then \(x+y = (\alpha+\beta)p - (u+v) \) with \(u+v \in [\![0,p-1]\!] \).
We have \(f(p-u) = f(p-v) \leqslant k\) and \(p-u + p-v > p\), the extra condition on \(f\) provides \(f(p-u-v) \neq
f(p-u)\).
We assumed \(h(x+y) \leqslant k\) hence \(h(x+y) = f(p-u-v)\). Since \(h(x) = f(p-u)\), we have at last \(h(x+y) \neq
h(x)\).
\item If \(u+v \geqslant p\), then \(x+y = (\alpha+\beta-1)p - (u+v-p) \) with \(u+v-p \in [\![0,p-1]\!] \).
We assumed \(h(x+y) \leqslant k \) hence \(h(x+y) = f( p- (u+v-p)) = f(p-u + p-v)\). Since \(f(p-u+p-v) \neq f(p-u)\),
we get \(h(x+y) \neq h(x)\).
\end{itemize}

We now assume \(h(x) > k\). Then \(h(x) = k + g(\alpha) = k + g(\beta) = h(y)\), hence \(g(\alpha) = g(\beta)\). 
We have the two same cases as before.

\begin{itemize}
\item If \(u+v < p\), then \(x+y = (\alpha+\beta)p - (u+v) \) with \(u+v \in [\![0,p-1]\!] \).
We assume \(h(x+y) > k\) otherwise the expected result is trivial.
Since \(g(\alpha) = g(\beta)\), the sum-free condition garanties \(g(\alpha) \neq g(\alpha+\beta)\), 
thus \(h(x+y) = k + g(\alpha + \beta) \neq k + g(\alpha) = h(x)\).
\item If \(u+v \geqslant p\), then \(x+y = (\alpha+\beta-1)p - (u+v-p) \) with \(u+v-p \in [\![0,p-1]\!] \).
Because of the assumption \(h(x) > k\), we necessarily have \(f(p-u) = k+1 = f(p-v)\). 
Then \(f(p-(u+v-p)) = f(p-u + p-v) \leqslant k \) with the sum-free condition on \(f\), 
hence \(h(x+y) = f(p-(u+v-p)) \) by construction of \(h\). Thus \(h(x+y) \leqslant k < h(x)\).
\end{itemize}
\end{proof}

@@

\par
Here are the best inequalities on Schur numbers so far (the templates corresponding to the third, fourth and fifth 
inequalities can be found in the appendix):
\[ S(n+1) \geqslant 3S(n) + 1 \]
\[ S(n+2) \geqslant 9S(n) + 4 \]
\[ S(n+3) \geqslant 33S(n) + 6 \]
\[ S(n+4) \geqslant 111S(n) + 43 \]
\[ S(n+5) \geqslant 380S(n) + 148 \]
\[ S(n+6) \geqslant 1140S(n) + 528 \]

>>>>

\begin{proof}[\textsc{Proof.}]
Denote by \(f\) the colouring associated to the \(SF\)-template of lenght \(p\) and \(g\) the one associated 
to the sum-free partition of \([\![1,q]\!]\); where \(f : [\![1,p]\!] \longrightarrow [\![1,k+1]\!]\) and 
\(g : [\![1,q]\!] \longrightarrow [\![1,n]\!]\).

\par
The sum-free condition is expressed as
\[\forall (x,y) \in [\![1,p]\!]^2, f(x) = f(y) \Longrightarrow f(x+y) \neq f(x)\]
\[\forall (x,y) \in [\![1,q]\!]^2 \text{, } g(x) = g(y) \Longrightarrow g(x+y) \neq g(x)\]

The additionnal constraint for the SF-template is
\[
\forall (x,y) \in [\![1,p]\!]^2, \left\{
\begin{array}{ll}
	f(x) = f(y) \leqslant k \\
	x + y > p
\end{array}
\right. \Longrightarrow f(x+y-p) \neq f(x)
\]

For \(x \in [\![1,pq+m_{k+1}-1]\!]\), write \(x = (\alpha - 1) + u\) for some integers \(\alpha \in \mathbb{Z}\) and 
\(u \in [\![1,p]\!]\). This decomposition is of course unique. \(\alpha\) (resp. \(u\)) can be interpreted as the row 
(resp. column) number of \(x\). Define a new coloring \(h\) as follows
\[
\begin{array}{c c c l}
	h : & [\![1,pq+m_{k+1}-1]\!] & \longrightarrow & [\![1,n+k]\!] \\
	& x & \longmapsto & 
	\left\{ \begin{array}{l l}
		f(u) & \text{if}~f(u) \leqslant k \\
		k + g(\alpha) & \text{if}~f(u) = k + 1 \\
	\end{array} \right.
\end{array}
\]

\(h\) is well-defined since, by definition of \(m_{k+1}\), \(\forall x \in [\![p q + 1, p q + m_{k + 1} - 1 ]\!], f(u) 
\leqslant k\) and therefore \(\forall x \in [\![1,pq+m_{k+1}-1]\!], f(u) = k + 1 \implies \alpha \in [\![1, q]\!]\).

\par
We now prove that \(h\) is a sum-free coloring. Let \(x,y \in [\![1,pq + m_{k+1}-1]\!]\) such that \(h(x) = h(y)\) 
and \(x+y \leqslant pq+m_{k+1}-1\). We claim that \(h(x+y) \neq h(x)\). Write \(x = (\alpha - 1) p + u\) and 
\(y = (\beta - 1) p + v\) where \(\alpha, \beta \in \mathbb{Z}\) and \(u, v \in [\![1,p]\!]\). Distinguish between 
two cases according to the value of \(h(x)\). \\

\noindent \underline{\textbf{First case:} \(h(x) \leqslant k\)} \\
\qquad Assume that \(h(x+y) \leqslant k\), otherwise \(h(x + y) \neq h(x)\) obviously holds. By definition of \(h\) 
and given that \(h(u) = h(v)\), \(f(u) = f(v)\). Distinguish between two cases according to the value of \(x + y\).

\begin{itemize}
\item If \(u + v > p\), write \(w = u + v - p \in [\![1, p]\!]\). Then \(x + y = (\alpha + \beta - 1) p + w\). By definition, 
	\(h(x + y) = f(w)\). Given that \(f(u) = f(v) \leqslant k\), the additionnal constraint on \(f\) implies \(f(w) 
	\neq f(u)\), that is \(h(x + y) \neq h(x)\).
\item If \(u + v \leqslant p\), write \(w = u + v \in [\![1, p]\!]\). Then \(x+y = (\alpha + \beta- 2) p + w\). By definition, 
	\(h(x + y) = f(w)\). Given that \(f(u) = f(v) \leqslant k\), the sum-free property of \(f\) implies \(f(w) \neq f(u)\), 
	that is \(h(x + y) \neq h(x)\).
\end{itemize}

\noindent \underline{\textbf{Second case:} \(h(x) \geqslant k + 1\)} \\
\qquad \(h(x) = k + g(\alpha) = 
k + g(\beta) = h(y)\), hence \(g(\alpha) = g(\beta)\). As in the first case, distinguish between two cases according 
to the value of \(x + y\).

\begin{itemize}
\item If \(u + v > p\), write \(w = u + v - p \in [\![1, p]\!]\). Then \(x + y = (\alpha + \beta - 1) p + w\). Assume that 
	\(h(x+y) \geqslant k + 1\),  otherwise \(h(x + y) \neq h(x)\) obviously holds. By definition, \(h(x + y) = k + g(\alpha + 
	\beta)\). Given that \(g(\alpha) = g(\beta)\), the sum-free property of \(g\) implies \(g(\alpha + \beta) 
	\neq g(\alpha)\) that is \(h(x + y) \neq h(x)\).
\item  If \(u + v \leqslant p\), write \(w = u + v \in [\![1, p]\!]\). Then \(x+y = (\alpha + \beta- 2) p + w\). The sum-free 
	property of \(f\) implies \(f(w) \neq f(u)\). Therefore \(f(w) \leqslant k\) and thus \(h(x + y) \leqslant k\). In particular,
	given that \(h(x) \geqslant k + 1\), \(h(x + y) \neq h(x)\).
\end{itemize}
\end{proof}

@@

\par
Here are the best inequalities on Schur numbers so far (the templates corresponding to the third, fourth and fifth 
inequalities can be found in the appendix):
\[
\begin{array}{r r r r r}
	S(n + 1) & \geqslant & 3 S(n) & + & 1 \\
	S(n + 2) & \geqslant & 9 S(n) & + & 4 \\
	S(n + 3) & \geqslant & 33 S(n) & + & 6 \\
	S(n + 4) & \geqslant & 111 S(n) & + & 43 \\
	S(n + 5) & \geqslant & 380 S(n) & + & 148 \\
	S(n + 6) & \geqslant & 1140 S(n) & + & 528 \\
\end{array}
\]

### 2021-05-02 article/weakschur.tex
modfis demos

\begin{corollary}
\( \forall (n,k) \in (\mathbb{N}^*)^2 \text{, } \WS (n+k) \geqslant S(k) \left (\WS (n) + \left \lceil \frac{\WS (n)}{2}
\right \rceil +1 \right) + \WS (n)\)
\end{corollary}

@@

Like for SF-templates, \(\mathcal{R}\) and \(\mathcal{C}\) play the role of security zones for each other. Note that with
this numbering of columns, the column of the sum of two numbers is the only integer in \([\![-l,q]\!]\) equal to two the
sum of the columns modulo \(a\).

\renewcommand{\arraystretch}{2}
\begin{center}
\[
\begin{array}{|c|c|c|c|c|c|c|c|c|>{\columncolor{blue}}c|>{\columncolor{blue}}c|>{\columncolor{blue}}c|>{\columncolor{green}} c|}
\cline{6-9}
   \multicolumn{5}{c|}{} & \cellcolor{green} 1 & \cellcolor{green} 2 & \cellcolor{blue} ... &\cellcolor{green} r & r + 1 & ... & b - 1 &  b \\
\hline
   \cellcolor{red} a - l & \cellcolor{red} a- l + 1 & \cellcolor{red} ... & \cellcolor{red} a - 1 & \cellcolor{red} a & \cellcolor{red} a + 1 & \cellcolor{red} ... &\cellcolor{red} a + r - 1 & \cellcolor{red} a + r & a + r + 1 & ... & a + b - 1 & a+ b \\
\hline
\cellcolor{yellow} 2 a - l &\cellcolor{yellow}...&\cellcolor{yellow}...&\cellcolor{yellow}...&\cellcolor{yellow} 2 a &\cellcolor{yellow} ... &\cellcolor{yellow} ... &\cellcolor{yellow} ... &\cellcolor{yellow} 2 a + r & ... & ... & ... & 2 a + b \\
\hline
...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...&...&...&...\\
\hline
\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&...&...&...&...\\
\hline
p a - l & ... & ... & ... &p a & ... & ... & .... & p a + r & ... & ... & ... & p a + b\\
\hline
\end{array}
\]
\end{center}

@@

In particular, taking \(w = l = \lceil \frac{q}{2} \rceil\) and \(r = \lfloor \frac{q}{2} \rfloor\) works, thus obtaining the above theorem.\\
\par
As in the previous section, we now introduce WSF-templates and the sequence \(\WS^+\) in order to generalize the above construction.

@@

\qquad In this subsection, we introduce WSF-templates and define the objects and prove the results needed
for the general theorem regarding templates for weak Schur numbers.

@@

Let \((a,b) \in (\mathbb{N}^*)^2\),\(a>b\), \(n \in \mathbb{N}^*\), we will define \(\pi_{a,b}\) the projection:
\[ \pi_{a,b}:x \longmapsto (Id+a\mathds{1}_{ [\![0,b]\!]})(x \mod a))\]
\end{definition}

We will note  the projection \(\pi\) and not \(\pi_{a,b}\) when there is no doubt about the \(a\) and \(b\) we use.

@@

\begin{proposition}
Let x \(\in [\![1,b]\!]\), let  \(y \in \mathbb{N}^*\) such that \(x+\pi(y)\leqslant a+b\), then we have: \(\pi(x+y)=x+\pi(y)\)
\end{proposition}

\begin{proof}[\textsc{Proof.}]
Let  \(x\in [\![1,b]\!]\), let  \(y\in \mathbb{N}^*\) such that \(x+\pi(y)\leqslant a+b\)

if \(x+\pi(y)< a:\)we remark that \(\pi(y)>b\) and therefore \(x+\pi(y)>b\):

\begin{align*}
 \pi(x+y) & = (Id+a\mathds{1}_{ [\![0,b]\!]})(x+y \mod a)\\
& = (Id+a\mathds{1}_{ [\![0,b]\!]})(x+\pi(y) \mod a)\text{ since } \pi(y)=y \mod a \\
& =x+ \pi(y)
\end{align*}

if \(x+\pi(y)\geqslant a\):
\begin{align*}
 \pi(x+y) & = (Id+a\mathds{1}_{ [\![0,b]\!]})(x+y \mod a)\\
& = (Id+a\mathds{1}_{ [\![0,b]\!]})(x+\pi(y)\mod a)\text{ since } \pi(y)=y \mod a \\
& = (Id+a\mathds{1}_{ [\![0,b]\!]})(x+\pi(y) - a) \\
& = x+\pi(y) - a +a\mathds{1}_{ [\![0,b]\!]})(x+\pi(y) - a) \\
& = x+\pi(y) - a +a \text{ \quad since } x+\pi(y) - a \in [\![0,b]\!] \\
& = x+\pi(y)
\end{align*}
\end{proof}

\begin{proposition}
Let \((x,y)\in (\mathbb{N}^*)^2\), \(\pi(\pi(x)+\pi(y))=\pi(x+y)\)
\end{proposition}

\begin{proof}[\textsc{Proof.}]
Let \((x,y)\in (\mathbb{N}^*)^2\),

\begin{align*}
 \pi(\pi(x)+\pi(y)) & = (Id+a\mathds{1}_{ [\![0,b]\!]})(\pi(x)+\pi(y) \mod a)\\
& = (Id+a\mathds{1}_{ [\![0,b]\!]})((Id+a\mathds{1}_{ [\![0,b]\!]})(x \mod a)+(Id+a\mathds{1}_{ [\![0,b]\!]})(y \mod a)\ mod a) \\
& = (Id+a\mathds{1}_{ [\![0,b]\!]})(((x \mod a) + (y \mod a)) \mod a)\\
& =(Id+a\mathds{1}_{ [\![0,b]\!]})( x+y \mod a)\\
& =\pi(x+y)
\end{align*}
\end{proof}

@@

Let \((a,b) \in (\mathbb{N}^*)^2\), \(a>b\), let \(n \in\mathbb{N}\) we will define \(\lambda_{a,b}\) the projection:
\[ \lambda_{a,b}:x \longmapsto 1+ \left\lfloor\dfrac{x-b-1}{a}\right\rfloor\]
\end{definition}

We will note  the projection \(\lambda\) and not \(\lambda_{a,b}\) when there is no doubt about the \(a\) and \(b\) we use.

In the following theorem, \(\lambda\) is the function which maps an element x to its line number as mentioned in the intuitive explanation.

\begin{proposition}
Let \((a,b)\in (\mathbb{N}^*)^2\), \(a>b\), let  \(x\in \mathbb{N}^*\), \(x=a\lambda(x)+\pi(x)-a\)
\end{proposition}
\begin{proof}[\textsc{Proof.}]
Let \((a,b)\in (\mathbb{N}^*)^2\), \(a>b\), let  \(x\in \mathbb{N}^*\),\\
\(a\lambda(x)+\pi(x)-a=a\left\lfloor\dfrac{x-b-1}{a}\right\rfloor+(x \mod a)+\mathds{1}_{ [\![0,b]\!]}(x \mod a)\)

if \(x \mod a>b\): \(a\lambda(x)+\pi(x)-a=a\left\lfloor\dfrac{x}{a}\right\rfloor+x \mod a=x\)

if \(x \mod a \leqslant b\):
\(a\lambda(x)+\pi(x)-a=a \left( \left \lfloor \dfrac{x}{a} \right \rfloor - 1 \right)+\text{x mod a}+a=x\)
\end{proof}

\begin{definition}
Let \( (p,n,b) \in (\mathbb{N}^*)^3\), Let \((A_1,...,A_n)\) a partition of  \([\![1, p]\!]\).
This partition is said to be a b-weakly-sum-free template (b-WSF-template) of \(n\) colors and lenght \(p\) when:
\\\\
\underline{\(\forall i \in [\![1, n]\!], \quad A_i\) is weakly-sum-free}
\\\\
\underline{\(\forall i \in [\![1, n]\!], \quad A_i\backslash [\![1, b]\!]\) is sum-free}
\\\\
\underline{For \(A_n\) (the special subset):} \quad \(\forall (x,y) \in A_n^2,\)
\\
\[if \quad x+y>b+2(p-b),\quad x+y-2(p-b)\notin A_n\]
\\
\underline{For the others subsets:}\quad \(\forall i \in [\![0,n-1]\!], \forall(x,y) \in A_i^2,\)
\\
\[
x+y>p \implies \pi(x+y) \notin A_i
\]
\end{definition}

>>>>

\begin{corollary}
\( \forall (n,k) \in (\mathbb{N}^*)^2 \text{, } \WS (n+k) \geqslant S(k) \left (\WS (n) + \left \lceil \displaystyle \frac{\WS (n)}{2}
\right \rceil +1 \right) + \WS (n)\)
\end{corollary}

@@

Like SF-templates, \(\mathcal{R}\) and \(\mathcal{C}\) play the role of security zones for each other. Note that with
this numbering of columns, the column of the sum of two numbers is the only integer in \([\![-l,q]\!]\) equal to two the
sum of the columns modulo \(a\).

\renewcommand{\arraystretch}{2}
\[
\begin{array}{|c|c|c|c|c|c|c|c|c|>{\columncolor{blue}}c|>{\columncolor{blue}}c|>{\columncolor{blue}}c|>{\columncolor{green}} c|}
\cline{6-9}
   \multicolumn{5}{c|}{} & \cellcolor{green} 1 & \cellcolor{green} 2 & \cellcolor{blue} ... &\cellcolor{green} r & r + 1 & ... & b - 1 &  b \\
\hline
   \cellcolor{red} a - l & \cellcolor{red} a- l + 1 & \cellcolor{red} ... & \cellcolor{red} a - 1 & \cellcolor{red} a & \cellcolor{red} a + 1 & \cellcolor{red} ... &\cellcolor{red} a + r - 1 & \cellcolor{red} a + r & a + r + 1 & ... & a + b - 1 & a+ b \\
\hline
\cellcolor{yellow} 2 a - l &\cellcolor{yellow}...&\cellcolor{yellow}...&\cellcolor{yellow}...&\cellcolor{yellow} 2 a &\cellcolor{yellow} ... &\cellcolor{yellow} ... &\cellcolor{yellow} ... &\cellcolor{yellow} 2 a + r & ... & ... & ... & 2 a + b \\
\hline
...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...&...&...&...\\
\hline
\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&...&...&...&...\\
\hline
p a - l & ... & ... & ... &p a & ... & ... & .... & p a + r & ... & ... & ... & p a + b\\
\hline
\end{array}
\]
\renewcommand{\arraystretch}{1}

@@

In particular, taking \(w = l = \lceil \displaystyle \frac{q}{2} \rceil\) and \(r = \lfloor \displaystyle \frac{q}{2} \rfloor\) works, thus obtaining the above theorem.\\
\par
As in the previous section, we now introduce WSF-templates and the sequence \(\WS^+\) in order to generalize the above construction.

@@

\qquad In this subsection, we introduce WSF-templates and define the objects and prove the results needed
for the general theorem on templates for weak Schur numbers.

@@

Let \((a,b) \in (\mathbb{N}^*)^2\) such that \(a>b\). Define
\[ \pi_{a,b}:x \longmapsto (Id+a\mathds{1}_{ [\![0,b]\!]})(x \mod a))\]
\end{definition}

If there is no confusion on the \(a\) and \(b\) to use, \(\pi_{a, b}\) is denoted by \(\pi\).

@@

\begin{proposition}
Let x \(\in [\![1,b]\!]\) and \(y \in \mathbb{N}^*\) such that \(x+\pi(y) \leqslant a+b\). Then \(\pi(x+y)=x+\pi(y)\).
\end{proposition}

\begin{proof}[\textsc{Proof.}]
Let  \(x\in [\![1,b]\!]\), and  \(y\in \mathbb{N}^*\) such that \(x+\pi(y)\leqslant a+b\).

\noindent if \(x+\pi(y)< a \):

\[
\begin{array}{c c l l}
	\pi(x+y) & = & (Id+a\mathds{1}_{ [\![0,b]\!]})(x+y \mod a) & \\
	& = & (Id+a\mathds{1}_{ [\![0,b]\!]})(x+\pi(y) \mod a) & \text{since } \pi(y)=y \mod a \\
	& = & x+ \pi(y) & 
\end{array}
\]

\noindent if \(x+\pi(y)\geqslant a\):
\[
\begin{array}{c c l l}
	\pi(x+y) & = & (Id+a\mathds{1}_{ [\![0,b]\!]})(x+y \mod a)& \\
	& = & (Id+a\mathds{1}_{ [\![0,b]\!]})(x+\pi(y)\mod a) & \text{since } \pi(y)=y \mod a \\
	& = & (Id+a\mathds{1}_{ [\![0,b]\!]})(x+\pi(y) - a) & \\
	& = & x+\pi(y) - a +a\mathds{1}_{ [\![0,b]\!]}(x+\pi(y) - a) & \\
	& = & x+\pi(y) - a +a & \text{since } x+\pi(y) - a \in [\![0,b]\!] \\
	& = & x+\pi(y) &
\end{array}
\]
\end{proof}

\begin{proposition}
Let \((x,y)\in (\mathbb{N}^*)^2\). Then \(\pi(\pi(x)+\pi(y))=\pi(x+y)\).
\end{proposition}

\begin{proof}[\textsc{Proof.}]
Let \((x,y)\in (\mathbb{N}^*)^2\).

\[
\begin{array}{c c l}
	\pi(\pi(x)+\pi(y)) & = & (Id+a\mathds{1}_{ [\![0,b]\!]})(\pi(x)+\pi(y) \mod a)\\
	& = & (Id+a\mathds{1}_{ [\![0,b]\!]})((Id+a\mathds{1}_{ [\![0,b]\!]})(x \mod a)+(Id+a\mathds{1}_{ [\![0,b]\!]})(y \mod a) \mod a) \\
	& = & (Id+a\mathds{1}_{ [\![0,b]\!]})(((x \mod a) + (y \mod a)) \mod a)\\
	& = & (Id+a\mathds{1}_{ [\![0,b]\!]})( x+y \mod a)\\
	& = & \pi(x+y)
\end{array}
\]
\end{proof}

@@

Let \((a,b) \in (\mathbb{N}^*)^2\) such that \(a>b\). Define
\[ \lambda_{a,b}:x \longmapsto 1+ \left\lfloor\dfrac{x-b-1}{a}\right\rfloor\]
\end{definition}

If there is no confusion on the \(a\) and \(b\) to use, \(\lambda_{a, b}\) is denoted by \(\lambda\).

\(\lambda\) is the function which maps an element x to its line number as mentioned in the intuitive explanation.

\begin{proposition}
Let \((a,b)\in (\mathbb{N}^*)^2\) such that \(a>b\) and let \(x\in \mathbb{N}^*\). Then \(x=a\lambda(x)+\pi(x)-a\).
\end{proposition}

\begin{proof}[\textsc{Proof.}]
Let \((a,b)\in (\mathbb{N}^*)^2\) such that \(a>b\) and let \(x\in \mathbb{N}^*\). \\
\(a\lambda(x)+\pi(x)-a=a\left\lfloor\dfrac{x-b-1}{a}\right\rfloor+(x \mod a)+ a \mathds{1}_{ [\![0,b]\!]}(x \mod a)\)

\noindent  if \(x \mod a>b\) then \(a\lambda(x)+\pi(x)-a=a\left\lfloor\dfrac{x}{a}\right\rfloor+x \mod a=x\)

\noindent if \(x \mod a \leqslant b\) then
\(a\lambda(x)+\pi(x)-a=a \left( \left \lfloor \dfrac{x}{a} \right \rfloor - 1 \right)+x \mod a +a=x\)
\end{proof}

\begin{definition}
Let \( (p,n,b) \in (\mathbb{N}^*)^3\), Let \((A_1,...,A_n)\) a partition of  \([\![1, p]\!]\).
This partition is said to be a b-weakly-sum-free template (b-WSF-template) of \(n\) colors and lenght \(p\) when:
\\\\
\underline{\(\forall i \in [\![1, n]\!], \quad A_i\) is weakly-sum-free}
\\\\
\underline{\(\forall i \in [\![1, n]\!], \quad A_i\backslash [\![1, b]\!]\) is sum-free}
\\\\
\underline{For \(A_n\) (the special subset):} \quad \(\forall (x,y) \in A_n^2,\)
\\
\[x+y>b+2(p-b) \implies x+y-2(p-b)\notin A_n\]
\\
\underline{For the others subsets:}\quad \(\forall i \in [\![0,n-1]\!], \forall(x,y) \in A_i^2,\)
\\
\[
x+y>p \implies \pi(x+y) \notin A_i
\]
\end{definition}
